---
title: "Plasma gcg"
author: "MWS"
date: "2023-10-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(data.table)
library(here)
library(limma)
library(tidyverse)
library(ggrepel)
library(kableExtra)
library(knitr)
library(ggpubr)
library(rstatix)
library(gridExtra)
```

## 1. Load data
```{r}
# Read olink protein expression dataframe

npx_df <- fread(here("data/olink_10%missing_MinProb_imputed.csv")) # 10% missing samples and proteins were removed before minProb imputation
#dim(npx_df2)
```


```{r}
gcg <- npx_df[, c("eid", "gcg")]

# Read the phenotype dataframe
pheno_df = fread(here("data-raw/df_gcgr_variants.csv"))

# Read baseline MASLD dataframe
MASLD_df <- fread(here("data/baseline_MASLD.csv"))
MASLD_df$MASLD_bs <- as.factor(MASLD_df$MASLD_bs) #MASLD: ICD10+PDFF+alcohol, MASLD_bs: PDFF+alcohol
MASLD_df$MASLD_bs <- factor(MASLD_df$MASLD_bs, levels = c(0,1), labels = c("No", "Yes"))
```

## 2. Merge and prepare data
```{r}
df <- merge(gcg, pheno_df, by = "eid")
df <- merge(df, MASLD_df, by = "eid")

# include G40S homozygot in the hetrozygot group
df[G40S == 2, G40S := 1]

# Copy percent_liver_fat to %liver_fat
df$percent_liver_fat <- df$PDFF
```

```{r}
# Make BMI and percent_liver_fat dichotomized
df[BMI <= 25, BMI_dic := "No"]
df[BMI >= 30, BMI_dic := "Yes"]
```

```{r}
# rename T1D and T2D
setnames(df, "adj_probposs_t1dm", "Type1Diabetes")
setnames(df, "adj_probposs_t2dm", "Type2Diabetes")

# Make sure T1D and gestational are removed from controls in T2D
df$adj_dmstatus_bl <- as.numeric(df$adj_dmstatus_bl)
df$adj_dmstatus_bl <-
    factor(df$adj_dmstatus_bl,
        levels = c(0,1,2,3,4,5),
        labels = c(
            "unlikely_diabetes",
            "poss_gestational",
            "prob_t1d",
            "poss_t1d",
            "prob_t2d",
            "poss_t2d"))
df[adj_dmstatus_bl == "poss_gestational" |
       adj_dmstatus_bl == "prob_t1d" |
       adj_dmstatus_bl == "poss_t1d",
   Type2Diabetes := NA] 

df[adj_dmstatus_bl == "poss_gestational" |
       adj_dmstatus_bl == "prob_t2d" |
       adj_dmstatus_bl == "poss_t2d",
   Type1Diabetes := NA]

count(df, adj_dmstatus_bl, Type2Diabetes, Type1Diabetes)

df$T2D <- factor(df$Type2Diabetes, levels = c(0,1), labels = c("No", "Yes"))
```


Log2 Fold change
```{r}
mean_BMI <- df[, list(mean=mean(gcg)), by=BMI_dic][!is.na(BMI_dic)]
paste("gcg fold change for BMI is", mean_BMI$mean[2]-mean_BMI$mean[1])

mean_MASLD <- df[, list(mean=mean(gcg)), by=MASLD_bs][!is.na(MASLD_bs)]
paste("gcg fold change for MASLD is", mean_MASLD$mean[2]-mean_MASLD$mean[1])

mean_t2d <- df[, list(mean=mean(gcg)), by=T2D]
paste("gcg fold change for T2D is", mean_t2d$mean[2]-mean_t2d$mean[1])
```

## 3. Simple boxplots of gcg and BMI, diabetes, MASLD

```{r gcg~BMI, warning=FALSE}
df_BMI <- df %>% filter(!is.na(BMI_dic)) 

# sample size 
sample_size <- df_BMI %>% 
  group_by(BMI_dic) %>% 
  summarise(num = n())

df_BMI <- df_BMI %>% 
    left_join(sample_size) %>% 
    mutate(xaxis = paste0(BMI_dic, "\n", "n=", num))

t_test_BMI <- t_test(gcg ~ BMI_dic, data = df_BMI) %>%
    add_significance() %>%
    add_xy_position(x = "BMI_dic")

boxplot_BMI <- ggboxplot(
    data = df_BMI,
    x = "xaxis",
    y = "gcg",
    #outlier.shape = NA,
    fill = "BMI_dic",
    xlab = "Obesity",
    ylab = "Plasma gcg (npx)",
    ylim = c(-6, 8),
    legend.position = "none",
        axis.title.y = element_text(size = 10)
      ) +
      stat_pvalue_manual(data = t_test_BMI, label = "p.signif") +
    theme(legend.position = "none")+ #removes legend
    scale_fill_manual(values = c("white", "grey"))
```

```{r gcg~T2D}
df_T2D <- df %>% filter(!is.na(T2D))

# sample size 
sample_size = df_T2D %>% group_by(T2D) %>%  summarise(num=n())
df_T2D <- df_T2D %>% 
    left_join(sample_size) %>% 
    mutate(xaxis = paste0(T2D, "\n", "n=", num))


t_test_T2D <- t_test(gcg ~ T2D, data = df_T2D) %>%
    add_significance() %>%
    add_xy_position(x = "T2D")

boxplot_T2D <- ggboxplot(
    data = df_T2D,
    x = "xaxis",
    y = "gcg",
    #outlier.shape = NA,
    fill = "T2D",
    xlab = "Type 2 diabetes",
    ylab = "Plasma gcg (npx)",
    ylim = c(-6, 8),
    legend.position = "none",
        axis.title.y = element_text(size = 14)
      ) +
      stat_pvalue_manual(data = t_test_T2D, label = "p.signif") +
    theme(legend.position = "none")+ #removes legend
    scale_fill_manual(values = c("white", "grey"))

```

```{r gcg~MASLD}
df_MASLD <- df %>% filter(!is.na(MASLD_bs))

# sample size 
sample_size = df_MASLD %>% group_by(MASLD_bs) %>%  summarise(num=n())
df_MASLD <- df_MASLD %>% 
    left_join(sample_size) %>% 
    mutate(xaxis = paste0(MASLD_bs, "\n", "n=", num))

t_test_MASLD <- t_test(gcg ~ MASLD_bs, data = df_MASLD) %>%
    add_significance() %>%
    add_xy_position(x = "MASLD_bs")

boxplot_MASLD <- ggboxplot(
    data = df_MASLD,
    x = "xaxis",
    y = "gcg",
    #outlier.shape = NA,
    fill = "MASLD_bs",
    xlab = "MASLD",
    ylim = c(-6, 8),
    ylab = "Plasma gcg (npx)",
    legend.position = "none",
        axis.title.y = element_text(size = 14)
      ) +
      stat_pvalue_manual(data = t_test_MASLD, label = "p.signif") +
    theme(legend.position = "none")+ #removes legend
    scale_fill_manual(values = c("white", "grey"))

```


## 4. Linear models adjusted for each of the other variables
```{r}
# Remove "alcoholics" from percent_liver_fat column
df[alcoholic == 1, percent_liver_fat := NA]
```

```{r}
# Adjust x-axis for BMI and PDFF for a better plot: effect of an increase of 5 of BMI/PDFF on plasma gcg
df$BMI_5units <- df$BMI/5
df$liver_fat_5units <- df$percent_liver_fat/5
```

###  4.1 BMI, T2D, PDFF
```{r Create_function_3_variables_adjust_for_each_other}
multiple_lm_threeVariables_adjust_each_other <- function(data, variable1, variable2, variable3) {
  # Convert data.table to data frame
  data_df <- as.data.frame(data)

  # Exclude rows with missing data in any of the specified variables
  data_df <- na.omit(data_df[, c("gcg", variable1, "age", "sex", "fasting_time", "pCreatinine", variable2, variable3)])
  
  # Count number of individuals included in the models
  n_individuals <- nrow(data_df)
  
  # Create the formula with variable names
  formula1 <- as.formula(paste("gcg ~", variable1, "+ age + sex + fasting_time + pCreatinine"))
  formula2 <- as.formula(paste("gcg ~", variable1, "+ age + sex + fasting_time + pCreatinine +", variable2))
  formula3 <- as.formula(paste("gcg ~", variable1, "+ age + sex + fasting_time + pCreatinine +", variable3))
  formula4 <- as.formula(paste("gcg ~", variable1, "+ age + sex + fasting_time + pCreatinine +", variable2, "+", variable3))

  # Model 1
  model1 <- lm(formula1, data = data_df)
  
  # Model 2
  model2 <- lm(formula2, data = data_df)
  
  # Model 3
  model3 <- lm(formula3, data = data_df)
  
  # Model 4
  model4 <- lm(formula4, data = data_df)
  
  # Extract estimates and p-values for 'variable1'
  results <- data.frame(
    Predictor = variable1,
    Model = c("Model 1", "Model 2", "Model 3", "Model 4"),
    Model1_additions = c("", variable2, variable3, paste(variable2, "+", variable3)),
    Estimate = c(coef(model1)[variable1], coef(model2)[variable1], coef(model3)[variable1], coef(model4)[variable1]),
    SE = c(summary(model1)$coef[variable1, "Std. Error"], summary(model2)$coef[variable1, "Std. Error"], 
                summary(model3)$coef[variable1, "Std. Error"], summary(model4)$coef[variable1, "Std. Error"]),
    P_Value = c(summary(model1)$coef[variable1, "Pr(>|t|)"], summary(model2)$coef[variable1, "Pr(>|t|)"], 
                summary(model3)$coef[variable1, "Pr(>|t|)"], summary(model4)$coef[variable1, "Pr(>|t|)"]),
    N_individuals <- n_individuals
  )
  
 
  return(results)
}
```

```{r apply function 3 variables}
results_BMI_5units <- multiple_lm_threeVariables_adjust_each_other(df, "BMI_5units", "Type2Diabetes", "liver_fat_5units")
results_T2D <- multiple_lm_threeVariables_adjust_each_other(df, "Type2Diabetes", "BMI_5units", "liver_fat_5units")
results_liver_fat_5units <- multiple_lm_threeVariables_adjust_each_other(df, "liver_fat_5units", "BMI_5units", "Type2Diabetes")
results_combined <- rbind(results_BMI_5units, results_T2D, results_liver_fat_5units)
results_combined$Adj.p.val <- p.adjust(results_combined$P_Value, method = "fdr")

# # Make variable by merging Predictor + model + model.additions
setDT(results_combined)
results_combined[Model %in% c("Model 2", "Model 3", "Model 4"), 
                  Variable := paste0(Predictor, ": ", Model, " = Model 1 + ", Model1_additions)]
results_combined[Model %in% "Model 1", Variable := paste0(Predictor, ":", Model)]

# Rename BMI_5units to BMI and liver_fat_5units to liver_fat
  results_combined$Variable <- gsub("BMI_5units", "BMI", results_combined$Variable)
  results_combined$Variable <- gsub("liver_fat_5units", "Liver fat", results_combined$Variable)
  results_combined$Variable <- gsub("Type2Diabetes", "T2D", results_combined$Variable)
  
  results_combined$Predictor <- gsub("BMI_5units", "BMI", results_combined$Predictor)
  results_combined$Predictor <- gsub("liver_fat_5units", "Liver fat", results_combined$Predictor)
  results_combined$Predictor <- gsub("Type2Diabetes", "T2D", results_combined$Predictor)

  # Specify order in plot:
results_combined$Variable <- factor(results_combined$Variable, levels = rev(results_combined$Variable))



```

### 4.2 BMI, T2D
```{r Create_function_2_variables_adjust_for_each_other}
multiple_lm_twoVariables_adjust_each_other <- function(data, variable1, variable2) {
  # Convert data.table to data frame
  data_df <- as.data.frame(data)

  # Exclude rows with missing data in any of the specified variables
  data_df <- na.omit(data_df[, c("gcg", variable1, "age", "sex", "fasting_time", "pCreatinine", variable2)])
  
  # Count number of individuals included in the models
  n_individuals <- nrow(data_df)
  
  # Create the formula with variable names
  formula1 <- as.formula(paste("gcg ~", variable1, "+ age + sex + fasting_time + pCreatinine"))
  formula2 <- as.formula(paste("gcg ~", variable1, "+ age + sex + fasting_time + pCreatinine +", variable2))
  
  # Model 1
  model1 <- lm(formula1, data = data_df)
  
  # Model 2
  model2 <- lm(formula2, data = data_df)
  
   # Extract estimates and p-values for 'variable1'
  results <- data.frame(
    Predictor = variable1,
    Model = c("Model 1", "Model 2"),
    Model1_additions = c("", variable2),
    Estimate = c(coef(model1)[variable1], coef(model2)[variable1]),
    SE = c(summary(model1)$coef[variable1, "Std. Error"], summary(model2)$coef[variable1, "Std. Error"]),
    P_Value = c(summary(model1)$coef[variable1, "Pr(>|t|)"], summary(model2)$coef[variable1, "Pr(>|t|)"]),
    N_individuals <- n_individuals
  )
  
  return(results)
}
```

```{r apply function 2 variables}
results_BMI_5units <- multiple_lm_twoVariables_adjust_each_other(df, "BMI_5units", "Type2Diabetes")
results_T2D <- multiple_lm_twoVariables_adjust_each_other(df, "Type2Diabetes", "BMI_5units")

results_combined2 <- rbind(results_BMI_5units, results_T2D)
results_combined2$Adj.p.val <- p.adjust(results_combined2$P_Value, method = "fdr")

# Make variable by merging Predictor + model + model.additions
setDT(results_combined2)
results_combined2[Model %in% "Model 2", Variable := paste0(Predictor, ":", Model, " = Model 1 + ", Model1_additions)]
results_combined2[Model %in% "Model 1", Variable := paste0(Predictor, ":", Model)]

#Rename BMI_5units to BMI and liver_fat_5units to liver_fat
  results_combined2$Variable <- gsub("BMI_5units", "BMI", results_combined2$Variable)
  results_combined2$Variable <- gsub("liver_fat_5units", "Liver fat", results_combined2$Variable)
  results_combined2$Variable <- gsub("Type2Diabetes", "T2D", results_combined2$Variable)
  
  results_combined2$Predictor <- gsub("BMI_5units", "BMI", results_combined2$Predictor)
  results_combined2$Predictor <- gsub("liver_fat_5units", "Liver fat", results_combined2$Predictor)
  results_combined2$Predictor <- gsub("Type2Diabetes", "T2D", results_combined2$Predictor)
  
# Specify order in plot:
results_combined2$Variable <- factor(results_combined2$Variable, levels = rev(results_combined2$Variable))

relocate(results_combined2, Variable, .before = )
```


## 5.  Forest plots 
```{r plot-function}
generate_forest_plot <- function(coef_table, N) {
  
   # Create the forest plot
  p <- ggplot(coef_table, aes(x = Estimate, xmin = Estimate - 1.96 * SE, xmax = Estimate + 1.96 * SE, 
                              y = Variable)) + 
    geom_point(position = position_dodge(width = 0.5), size = 3) +
    geom_errorbarh(position = position_dodge(width = 0.5), height = 0) +
    scale_x_continuous() +
    labs(x = paste0("Change in plasma pro-glucagon per increment (95% CI) \n N = ", N), y = "") +
    geom_vline(xintercept = 0, linetype = "dashed", color = "gray") +
    theme_classic() +
    theme(legend.position = "none",
        strip.placement = "outside",  # Place the strip labels outside the plot
        strip.background = element_blank(),  # Remove strip background
        axis.line.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.text.y = element_text(size = 12),
        axis.text.x = element_text(size = 12),
        strip.text.y.left = element_text(face = "bold", angle = 0, size = 11)) +  # Set angle to 0 for horizontal text
    facet_grid(Predictor ~ ., scales = "free", space = "free", switch = "y") +
  theme(plot.margin = margin(l = 30))  # Adjust left margin to create space for strip labels)

  # Create the first table
  t1 <- ggplot(data = coef_table) +
    geom_text(aes(y = Variable, x = 1, label = paste0(round(Estimate, digits = 2))), vjust = 0) +
    ggtitle(expression(beta)) +
    xlab(" ") +
    theme_classic(base_size = 13) +
    theme(
      axis.line = element_blank(),
      axis.text.y = element_blank(),
      axis.ticks = element_blank(),
      axis.title.y = element_blank(),
      axis.text.x = element_text(color = "white"),
      plot.title = element_text(size = 12, hjust = 0.5)
    )

  # Create the second table
  t2 <- ggplot(data = coef_table) +
    geom_text(aes(y = Variable, x = 1, label = paste0("(", round(Estimate - 1.96 * SE, digits = 2), "; ", round(Estimate + 1.96 * SE, digits = 2), ")")), vjust = 0) +
    ggtitle("95% CI") +
    xlab("  ") +
    theme_classic(base_size = 13) +
    theme(
      axis.line = element_blank(),
      axis.text.y = element_blank(),
      axis.ticks = element_blank(),
      axis.title.y = element_blank(),
      axis.text.x = element_text(color = "white"),
      plot.title = element_text(size = 12, hjust = 0.5)
    )

  # Create the third table
  t3 <- ggplot(data = coef_table) +
    geom_text(aes(y = Variable, x = 1, label = signif(Adj.p.val, digits = 2)), vjust = 0) +
    ggtitle("p-value") +
    xlab("  ") +
    theme_classic(base_size = 13) +
    theme(
      axis.line = element_blank(),
      axis.text.y = element_blank(),
      axis.ticks = element_blank(),
      axis.title.y = element_blank(),
      axis.text.x = element_text(color = "white"),
      plot.title = element_text(size = 12, hjust = 0.5)
    )

  # Arrange the plots and tables using grid.arrange
  combined_plot <- grid.arrange(p, t1, t2, t3, widths = c(15, 1, 2, 2))
  
  return(combined_plot)
}

```

```{r}
p1 <- generate_forest_plot(results_combined, "2,857")

```

```{r}
p2 <- generate_forest_plot(results_combined2, "37,979")

```

########################################################################################################################

## 6. Linear models strata on genotype (functions)
First we look at the effect of gcg on diseases seperately for the three genotypes
```{r create_function_strata_dic}
calculate_interaction_coefficients_glm <- function(data, genotype, outcome) {
 
  # Filter out missing values for genotype and outcome
  df_filtered <- data %>%
    filter(!is.na({{genotype}}), !is.na(.data[[outcome]]))

  # Calculate the sample size for each level of genotype and disease
  sample_sizes <- df_filtered %>%
    group_by({{genotype}}, .data[[outcome]]) %>%
    summarise(sample_size = n())

  # Perform the glm model separately for each level of genotype
  models <- df_filtered %>%
    group_by({{genotype}}) %>%
    do(model = glm(paste({{outcome}}, "~ gcg+age+sex+BMI+fasting_time+pCreatinine"), data = ., family = "binomial")) %>%
    pull(model)

  # Summarize the models and extract coefficients with SE
  coef_summaries <- lapply(models, function(model) {
    coef_summary <- summary(model)$coefficients
    data.frame(
      Estimate = coef_summary["gcg", "Estimate"],
      SE = coef_summary["gcg", "Std. Error"],
      P.value = coef_summary["gcg", "Pr(>|z|)"]
    )
  })

  # Calculate 95% CI from SE
  coef_summaries <- lapply(coef_summaries, function(coef_summary) {
    coef_summary$CI <- paste0(
      "(", round(coef_summary$Estimate - 1.96 * coef_summary$SE, 2),
      ", ", round(coef_summary$Estimate + 1.96 * coef_summary$SE, 2), ")"
    )
    return(coef_summary)
  })

  # Combine the coefficient summaries with sample sizes
  coef_table <- data.frame(
    variable = outcome,
    sample_size_wt_ctr = sample_sizes$sample_size[1],
    sample_size_wt_disease = sample_sizes$sample_size[2],
    sample_size_genotype_ctr = sample_sizes$sample_size[3],
    sample_size_genotype_disease = sample_sizes$sample_size[4],
    Estimate_wt = coef_summaries[[1]]$Estimate,
    CI_wt = coef_summaries[[1]]$CI,
    P.value_wt = coef_summaries[[1]]$P.value,
    Estimate_genotype = coef_summaries[[2]]$Estimate,
    CI_genotype = coef_summaries[[2]]$CI,
    P.value_genotype = coef_summaries[[2]]$P.value
  )

  return(coef_table)
}

```

```{r create_function_strata_con}
calculate_interaction_coefficients_lm <- function(data, genotype, outcome) {

  
  # Filter out missing values for genotype and the variable of interest
  df_filtered <- data %>%
    filter(!is.na({{genotype}}), !is.na(.data[[outcome]]))

  # Calculate the sample size for each level of genotype
  sample_sizes <- df_filtered %>%
    group_by({{genotype}}) %>%
    summarise(sample_size = n())

  # Perform the linear regression model separately for each level of genotype
  models <- df_filtered %>%
    group_by({{genotype}}) %>%
    do(model = lm(paste(outcome, "~ gcg+age+sex+Type2Diabetes+fasting_time+pCreatinine"), data = .)) %>%
    pull(model)

  # Summarize the models and extract coefficients
  coef_summaries <- lapply(models, function(model) {
    coef_summary <- summary(model)$coefficients
    ci <- confint(model)["gcg", ]
    data.frame(
      Estimate = coef_summary["gcg", "Estimate"],
      CI = paste0("(", round(ci[1], 2), ", ", round(ci[2], 2), ")"),
      P.value = coef_summary["gcg", "Pr(>|t|)"]
    )
  })

  # Combine the coefficient summaries with sample sizes
  coef_sum_df <- data.frame(
    variable = outcome,
    sample_size_wt_ctr = sample_sizes$sample_size[1],
    sample_size_wt_disease = NA,
    sample_size_genotype_ctr = sample_sizes$sample_size[2],
    sample_size_genotype_disease = NA,
    Estimate_wt = coef_summaries[[1]]$Estimate,
    CI_wt = coef_summaries[[1]]$CI,
    P.value_wt = coef_summaries[[1]]$P.value,
    Estimate_genotype = coef_summaries[[2]]$Estimate,
    CI_genotype = coef_summaries[[2]]$CI,
    P.value_genotype = coef_summaries[[2]]$P.value
  )

  return(coef_sum_df)
}
```


```{r}
# Apply the two functions
gcg_gcgr_interaction_cAMP <- calculate_interaction_coefficients_glm(df, cAMP_LoF, "Type2Diabetes")
gcg_gcgr_interaction_trunc <- calculate_interaction_coefficients_glm(df, trunc, "Type2Diabetes")
gcg_gcgr_interaction_G40S <- calculate_interaction_coefficients_glm(df, G40S, "Type2Diabetes")

gcg_gcgr_interaction_cAMP2 <- calculate_interaction_coefficients_lm(df, cAMP_LoF, "BMI")
gcg_gcgr_interaction_trunc2 <- calculate_interaction_coefficients_lm(df, trunc, "BMI")
gcg_gcgr_interaction_G40S2 <- calculate_interaction_coefficients_lm(df, G40S, "BMI")

gcg_gcgr_interaction_cAMP <- rbind(gcg_gcgr_interaction_cAMP, gcg_gcgr_interaction_cAMP2)
gcg_gcgr_interaction_trunc <- rbind(gcg_gcgr_interaction_trunc, gcg_gcgr_interaction_trunc2)
gcg_gcgr_interaction_G40S <- rbind(gcg_gcgr_interaction_G40S, gcg_gcgr_interaction_G40S2)
```

#############################################################################################################

## 7. Test for interaction between gcg and GCGR on phenotype
The p-value for this analysis show whether the effect for gcg in a specific genotype vs. ctr (above in 6.) is significantly different. 
```{r}
# cAMP - t2d
summary(glm(Type2Diabetes ~ gcg*cAMP_LoF+age+sex+BMI+fasting_time+pCreatinine, data = df, family = "binomial"))
 # gcg:cAMP_LoF P = 0.8

# cAMP - BMI
summary(lm(BMI ~ gcg*cAMP_LoF+age+sex+BMI+fasting_time+pCreatinine, data = df))
 # gcg:cAMP_LoF P = 0.1

# =========================================

# trunc - t2d
summary(glm(Type2Diabetes ~ gcg*trunc+age+sex+BMI+fasting_time+pCreatinine, data = df, family = "binomial"))
 # gcg:trunc P = 0.99
# trunc - BMI
summary(lm(BMI ~ gcg*trunc+age+sex+BMI+fasting_time+pCreatinine, data = df))
 # gcg:trunc P = 0.4

# ========================================

# G40S - t2d
summary(glm(Type2Diabetes ~ gcg*G40S+age+sex+BMI+fasting_time+pCreatinine, data = df, family = "binomial"))
 # gcg:G40S P = 0.07.

# G40S - BMI
summary(lm(BMI ~ gcg*G40S+age+sex+BMI+fasting_time+pCreatinine, data = df))
 # gcg:G40S P = 0.01*
```


