---
title: "Cox analysis gcg t2d"
author: "MWS"
date: "2023-11-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(data.table)
library(here)
library(tidyverse)
library(zoo)
library(survival)
library(survminer)
library(Hmisc)
library(gridExtra)
library(openxlsx)
source(here("R/functions.R"))
```

## 1. Import data and rename variables

```{r}
# Read the phenotype dataframe
df = fread("~/Gcgr_variants/data-raw/ukb_all_datasets.csv")

# Remove withdrawals
withdrawals <- fread(here("data-raw/xxx.txt"))
df <- df[!df$eid %in% withdrawals$V1,]

# Remove outdated collumns for dates for E11 and E14 diagnoses - these were downloaded from DNA nexus seperately as they were not in the original UKB data basket that was updated 1st oct 2023
df[, date_e11_first_reported_noninsulindependent_diabetes_mellitus_f130708_0_0 := NULL]
df[, date_e14_first_reported_unspecified_diabetes_mellitus_f130714_0_0 := NULL]

# Import updated columns for date for E11 and E14.
# (downloaded 15.11.23)
dates <- fread(here("data/updated_t2d_dates.csv"), na.strings = c("", "NA")) 
setnames(dates, "Participant ID", "eid")
df <- merge(df, dates, by = "eid", all.x = TRUE)
```

Calculate time between enrolment visit (visit0) and imaging visit (visit2) - used in article discussion
```{r}
time_to_visit2 <- as.yearmon(df$date_of_attending_assessment_centre_f53_2_0) - as.yearmon(df$date_of_attending_assessment_centre_f53_0_0)

median(time_to_visit2, na.rm = TRUE)

head(df$date_of_attending_assessment_centre_f53_2_0, 20)
head(df$date_of_attending_assessment_centre_f53_0_0, 20)
```


Create new dataset with only individuals that have olink data 
```{r}
npx_df <- fread(here("data/olink_10%missing_MinProb_imputed.csv")) # 10% missing samples and proteins were removed before minProb imputation

gcg <- npx_df[, c("eid", "gcg")]

#default for all = FALSE, i.e. remove rows that do not match on eid
df2 <- merge(df, gcg, by = "eid")
```


```{r}
# Change col names to shorter name
setnames(df2, "age_when_attended_assessment_centre_f21003_0_0", "age")
setnames(df2, "sex_f31_0_0", "sex")
setnames(df2, "fasting_time_f74_0_0", "fasting_time")
setnames(df2, "body_mass_index_bmi_f21001_0_0", "BMI")
setnames(df2, "glycated_haemoglobin_hba1c_f30750_0_0", "hba1c")
setnames(df2, "Date E11 first reported (non-insulin-dependent diabetes mellitus)", "date_e11_rep")
setnames(df2, "Date E14 first reported (unspecified diabetes mellitus)", "date_e14_rep")


```

```{r}

# For data downloaded from DNA nexus: "Code has event date matching participant's date of birth"
df2[date_e11_rep == "Code has event date matching participant's date of birth", date_e11_rep := NA]
df2$date_e11_rep <- as.IDate(df2$date_e11_rep) # was character string before because of the "matching date of birth"

```

```{r}
# Make categorical value for BMI
df2[BMI < 25, BMI_cat := "Normal"] # Underweight not included as there is only 1 female t2d case
df2[BMI >=25 & BMI < 30, BMI_cat := "Overweight"]
df2[BMI >= 30, BMI_cat := "Obese"]
df2[is.na(BMI), BMI_cat := NA]

df2$BMI_cat <- factor(df2$BMI_cat)

df2 %>% group_by(BMI_cat) %>%  summarise(mean(BMI, na.rm = TRUE), n())
```

## 2. Define T2D by ICD10

To calculate indicent T2D, we will use ICD codes that are updated continously (Set date and import fresh data from DNA nexus). Baseline T2D (Eastwood algorithm) is subtracted. 

```{r}
# T2D = E11 + E14
df2 <- find_ICD10_filter(df2, c("E11", "E14"), "T2D")

# T1D = E10
df2 <- find_ICD10_filter(df2, "E10", "T1D")

count(df2, T1D, T2D) # T2D = 3143
# Approx. 3/4 with T1D also have a code for T2D. 

df2[T1D == 1 & T2D == 1, T2D := 0]
count(df2, T2D) # T2D = 2816
```


## 4. Incident T2D, time to event

Exclude individuals with Eastwood algrithm for basleine T2D
```{r}
count(df2, T2D, adj_probposs_t2dm)
df2[T2D == 1 & adj_probposs_t2dm == 1, T2D := 0]
count(df2, T2D)
```


Find date for ICD10 code, calculate incident t2d
```{r}
# Merge the date columns and keep the earliest date of the two if both E11 and E14 are present
df3 <- df2 %>% 
   mutate(t2d_date = pmin(date_e11_rep, date_e14_rep))
df3[is.na(date_e11_rep), t2d_date := date_e14_rep]
df3[is.na(date_e14_rep), t2d_date := date_e11_rep]

# Create column of incident t2d 
df3[, incident_t2d := 0]
df3[T2D == 1 & t2d_date > date_of_attending_assessment_centre_f53_0_0, incident_t2d := 1]
count(df3, incident_t2d)
```

Exclude individuals with baseline HbA1c above 48 mmol/mol
```{r}
df3[hba1c > 48, incident_t2d := 0]
count(df3, incident_t2d)
```

Calculate time to event
```{r}
# set end date (when UKB data was last updated)
end_date <- as.Date("2021-10-03")

# Time til death (censored observations), units = months/years
df3$time_to_death <-
    as.numeric(as.yearmon(df3$date_of_death_f40000_0_0) - as.yearmon(df3$date_of_attending_assessment_centre_f53_0_0))

# Time to lost to follow-up (censored observations), units = months/years
df3$time_to_lost_to_followup <-
    as.numeric(as.yearmon(df3$date_lost_to_followup_f191_0_0) - as.yearmon(df3$date_of_attending_assessment_centre_f53_0_0))

# Time till t2d diagnosis, units = months/years, was defined above (for histogram)
# Remove diagnosis before baseline
df3[incident_t2d == 0, time_to_t2d := NA] 
count(df3, incident_t2d)

# Time from inclusion
df3$time_from_inclusion <- 
     as.numeric(as.yearmon(end_date) - as.yearmon(df3$date_of_attending_assessment_centre_f53_0_0))

# Median follow up time:
median(df3$time_from_inclusion) # 14.75

# Time to event. Event can be either death, lost to follow-up, or t2d diagnosis, or if NA in these, end_date
df3[incident_t2d == 1, time_to_event := time_to_t2d]
df3[incident_t2d == 0 & is.numeric(time_to_death), time_to_event := time_to_death]
df3[is.na(time_to_event) & incident_t2d == 0 & is.numeric(time_to_lost_to_followup), time_to_event := time_to_lost_to_followup]
df3[is.na(time_to_event), time_to_event := time_from_inclusion]


head(df3[, c("eid", "time_to_event", "date_of_death_f40000_0_0", "date_of_attending_assessment_centre_f53_0_0", "time_to_death", "t2d_date", "time_to_t2d", "incident_t2d")], 10)

df4 <- df3[, c("eid", "age", "sex", "BMI", "BMI_cat", "gcg", "time_to_event", "date_of_death_f40000_0_0", "date_of_attending_assessment_centre_f53_0_0", "time_from_inclusion", "time_to_death", "t2d_date", "time_to_t2d", "incident_t2d", "time_to_lost_to_followup")]
count(df3, incident_t2d)
```

```{r}
fwrite(df4, file = here("data/data_for_survival_analysis.csv"))
```


## 5. Cox analysis
### 5.1 All individuals

```{r}
df3 <- fread(here("data/data_for_survival_analysis.csv"))

df3[is.na(BMI), BMI_cat := NA] #otherwise it is just empty
df3$BMI_cat <- factor(df3$BMI_cat)
levels(df3$BMI_cat)

df3$age_5 <- df3$age/5
```

Remove incident cases within first 6 months after enrolment (=prevalent cases), and adjust baseline date by 6 months ( = subtract 0.5 years from time_to_event)

```{r}
# T2D diagnosis within the first 6 months after enrollment (diagnosis bias?)
df4 <- df3[!is.na(sex) & !is.na(BMI) & !is.na(age) & !is.na(gcg)]
df4[time_to_t2d < 0.5, incident_t2d := 0]
df4$time_to_event_adj <- df4$time_to_event-0.5
```

Create function to check linearity
```{r}
check_linearity = function(df, model, continuous_variable) {

  # Calculate martingale residuals
  mart_resid <- residuals(model, type = "martingale")

  # Plot MR vs. gcg
  plot(df[[continuous_variable]], mart_resid, xlab = paste(continuous_variable), ylab = "Martingale Residuals", main = paste("Martingale Residuals vs.", continuous_variable))

  # Add a smooth curve to visualize trends
  smoothed_line <- loess(mart_resid ~ df[[continuous_variable]])
  lines(df[[continuous_variable]], fitted(smoothed_line), col = "red")

   #Add a horizontal reference line at 0
  abline(h = 0, lty = 2)
}
```

## Cox models

Model 1
Kun age + sex
```{r}
model1 <- coxph(Surv(time_to_event_adj, incident_t2d) ~ gcg+age_5+sex, data = df4)
summary(model1)
cox.zph(model1)
```

Model 2
Age + sex + BMI continous
```{r}
model2 <- coxph(Surv(time_to_event_adj, incident_t2d) ~ gcg+age_5+sex+BMI, data = df4)
summary(model2)
cox.zph(model2)
```


Model 3
Both sexes, split on BMI in categories
```{r}
df_normal <- df4[BMI_cat == "Normal",]
df_overweight <- df4[BMI_cat == "Overweight",]
df_obese <- df4[BMI_cat == "Obese",]

model_normal <- coxph(Surv(time_to_event_adj, incident_t2d) ~ gcg+age_5+sex, data=df_normal)
summary(model_normal)
cox.zph(model_normal)
#check_linearity(df_normal, model_normal, "gcg")
#check_linearity(df_normal, model_normal, "age")

model_overweight <- coxph(Surv(time_to_event_adj, incident_t2d) ~ gcg+age_5+sex, data=df_overweight)
summary(model_overweight)
cox.zph(model_overweight)
#check_linearity(df_overweight, model_overweight, "gcg")
#check_linearity(df_overweight, model_overweight, "age")

model_obese <- coxph(Surv(time_to_event_adj, incident_t2d) ~ gcg+age_5+sex, data=df_obese)
summary(model_obese)
cox.zph(model_obese)
#check_linearity(df_obese, model_obese, "gcg")
#check_linearity(df_obese, model_obese, "age")

```


Model 4
Split on sex
```{r}
df_male <- df4[sex == "Male", ]
df_female <- df4[sex == "Female",]

model_male <- coxph(Surv(time_to_event_adj, incident_t2d) ~ gcg+age_5+BMI_cat, data=df_male)
summary(model_male)
cox.zph(model_male)


model_female <- coxph(Surv(time_to_event_adj, incident_t2d) ~ gcg+age_5+BMI_cat, data=df_female)
summary(model_female)
cox.zph(model_female)

model_male2 <- coxph(Surv(time_to_event_adj, incident_t2d) ~ gcg+age_5, data=df_male)
summary(model_male2)
cox.zph(model_male2)


model_female2 <- coxph(Surv(time_to_event_adj, incident_t2d) ~ gcg+age_5, data=df_female)
summary(model_female2)
cox.zph(model_female2)
```



Create function to create a table 
```{r}
cox_summary_table <- function(cox_model) {
  # Extract summary information
  summary_table <- summary(cox_model)

  # Extract coefficients and confidence intervals
  coefficients <- summary_table$coefficients[1, , drop = FALSE] # [1] because we only want to keep gcg which is the first covariate in he function
  ci <- confint(cox_model)[1, , drop = FALSE]
  
   # Extract number of events
  num_events <- summary_table$nevent

  # Create a data frame for the table
  result_table <- data.frame(
    BMI_group = deparse(substitute(cox_model)),  # Name of the cox_model, deparse(substitute(cox_model)) is used to obtain the name of the model as a string.
    covariate = rownames(coefficients), 
    HR = exp(coefficients[, "coef"]),
    lower_ci = exp(ci[, 1]),
    upper_ci = exp(ci[, 2]),
    p_value = coefficients[, "Pr(>|z|)"],
    num_events = num_events
  )

  # Replace variable names with more descriptive labels
  result_table$covariate <- gsub("sexMale", "sex (male)", result_table$covariate)
  result_table$covariate <- gsub("age_5", "age (5y)", result_table$covariate)

  # Round numeric values to 2 decimals
  result_table[, c("HR", "lower_ci", "upper_ci")] <- 
    round(result_table[, c("HR", "lower_ci", "upper_ci")], 2)

  # Format p-value in scientific notation with 2 decimal places
  result_table$p_value <- format(result_table$p_value, scientific = TRUE, digits = 2)

  return(result_table)
}
```

create coefficient table
```{r}
# Create table for plotting
result_table_normal <- cox_summary_table(model_normal)

result_table_overweight <- cox_summary_table(model_overweight)

result_table_obese <- cox_summary_table(model_obese)

combined_table <- rbind(result_table_normal, result_table_overweight, result_table_obese)

# change BMI group names
combined_table$BMI_group <- sub("model_normal", "BMI < 25", combined_table$BMI_group)
combined_table$BMI_group <- sub("model_overweight", "25 > BMI < 30", combined_table$BMI_group)
combined_table$BMI_group <- sub("model_obese", "BMI > 30", combined_table$BMI_group)
combined_table$BMI_group <- factor(combined_table$BMI_group, levels = c("BMI < 25", "25 > BMI < 30", "BMI > 30"))

write.xlsx(combined_table, file=here("data/cox_table_for_forestplot_BMIgroups.xlsx"))

```

Create forest plot
```{r}
p <- ggplot(combined_table, aes(x = HR, xmin = lower_ci, xmax = upper_ci, y = BMI_group)) +
  geom_point(position = position_dodge(width = 0.5), size=3) +
  geom_errorbarh(position = position_dodge(width = 0.5), height = 0) +
  scale_x_continuous(limits = c(0.9, 1.3)) + 
  labs(title = "", x = "HR (95% CI)", y = "") +
    geom_vline(xintercept = 1, linetype = "dashed", color = "gray") +  # Add a dashed line at x = 0
  theme_classic() +
    theme(legend.position = "left",
          axis.line.y = element_blank(),
          axis.ticks.y = element_blank(),
          axis.text.y = element_text(size = 12),
          axis.text.x = element_text(size = 12))


p
```

```{r tables_for_figure}
# Create the first table
  t1 <- ggplot(data = combined_table) +
    geom_text(aes(y = BMI_group, x = 1, label = HR), vjust = 0) +
    ggtitle("HR") +
    xlab(" ") +
    theme_classic(base_size = 13) +
    theme(
      axis.line = element_blank(),
      axis.text.y = element_blank(),
      axis.ticks = element_blank(),
      axis.title.y = element_blank(),
      axis.text.x = element_text(color = "white"),
      plot.title = element_text(size = 12, hjust = 0.5)
    )

 # Create the second table
  t2 <- ggplot(data = combined_table) +
    geom_text(aes(y = BMI_group, x = 1, 
                  label = paste0("(", lower_ci, "; ", upper_ci, ")")), 
              vjust = 0,
              position = position_dodge(width = 0.5)) +
    ggtitle("95% CI") +
    xlab("  ") +
    theme_classic(base_size = 13) +
    theme(
      axis.line = element_blank(),
      axis.text.y = element_blank(),
      axis.ticks = element_blank(),
      axis.title.y = element_blank(),
      axis.text.x = element_text(color = "white"),
      plot.title = element_text(size = 12, hjust = 0.5)
    )

# Create the third table
  t3 <- ggplot(data = combined_table) +
    geom_text(aes(y = BMI_group, x = 1, 
                  label = paste0(p_value)), 
              vjust = 0,
              position = position_dodge(width = 0.5)) +
    ggtitle("p-value") +
    xlab("  ") +
    theme_classic(base_size = 13) +
    theme(
      axis.line = element_blank(),
      axis.text.y = element_blank(),
      axis.ticks = element_blank(),
      axis.title.y = element_blank(),
      axis.text.x = element_text(color = "white"),
      plot.title = element_text(size = 12, hjust = 0.5)
    )
  
  # Create the fourth table
  t4 <- ggplot(data = combined_table) +
    geom_text(aes(y = BMI_group, x = 1, 
                  label = paste0(num_events)), 
              vjust = 0,
              position = position_dodge(width = 0.5)) +
    ggtitle("N cases") +
    xlab("  ") +
    theme_classic(base_size = 13) +
    theme(
      axis.line = element_blank(),
      axis.text.y = element_blank(),
      axis.ticks = element_blank(),
      axis.title.y = element_blank(),
      axis.text.x = element_text(color = "white"),
      plot.title = element_text(size = 12, hjust = 0.5)
    )
```

```{r combine_plot}
# Arrange the plots and tables using grid.arrange
 combined_plot <- grid.arrange(p, t1, t2, t3, t4, widths = c(7, 1, 1.5, 1, 1))

```




## 6. Kaplan-Meier plot
First create tertiles
```{r}
df4 <- fread(here("data/data_for_survival_analysis.csv"))
df4[time_to_t2d < 0.5, incident_t2d := 0]

df4$gcg_tertiles <- cut(df4$gcg, breaks = quantile(df4$gcg, probs = c(0, 1/3, 2/3, 1), na.rm = TRUE), labels = c("Tertile 1", "Tertile 2", "Tertile 3"))
relocate(df4, "gcg_tertiles", .after = "gcg")

df4 %>% group_by(gcg_tertiles) %>% summarise(mean = mean(gcg), n = n())

df4$gcg_quantiles <- cut(df4$gcg, breaks = quantile(df4$gcg, probs = c(0, 1/4, 2/4, 3/4, 1), na.rm = TRUE), labels = c("Quantile 1", "Quantile 2", "Quantile 3", "Quantile4"))
relocate(df4, "gcg_quantiles", .after = "gcg")

df4 %>% group_by(gcg_quantiles) %>% summarise(mean = mean(gcg), n = n())

```

```{r tertiles}
df4$sex <- factor(df4$sex, levels = c("Female", "Male"))

# Fit the survival model with 'gcg_tertiles'
fit <- survfit(Surv(time_to_event, incident_t2d) ~ gcg_tertiles, data = df4)

# Log-rank test overall
survdiff(Surv(time_to_event, incident_t2d) ~ gcg_tertiles, data = df4) #overall: significantly different, P<2e-16

# Pairwise log-rank tests with Bonferroni correction
pairwise_survdiff(Surv(time_to_event, incident_t2d) ~ gcg_tertiles, data = df4)
```


```{r tertiles}

# plot for both sexes
km_plot <- ggsurvplot(
    fit,
    df4,
    conf.int = TRUE,
    palette = "jco",
    ylim = c(0, 0.1),
    xlim = c(0, 17),
    ylab = "Risk of T2D",
    xlab = "Time (years)",
    censor = FALSE,
    legend.title = "Gcg tertile:  ",
    legend.labs = c("Low", "Medium", "High"),
    fun = "event" # Specify the function to plot the cumulative incidence
)

km_plot

```

