---
title: "obesity/NAFLD"
author: "MWS"
date: "2023-09-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(data.table)
library(tidyverse)
library(ggplot2)
library(MatchIt)
library(here)
library(epitools)
library(ggpubr)
library(rstatix)
library(logistf)
library(openxlsx)
library(gridExtra)
df <- fread(here("data/df_gcgr_variants.csv"))
```

# 1. Import and filter data

```{r}
count(df, trunc, cAMP_LoF, G40S, gcgr)

#Specify order of the levels (useful for later plotting)
df$gcgr <- factor(df$gcgr, c("WT", "G40S 0/1", "G40S 1/1", "cAMP_LoF 0/1", "truncated"))

# Import gcg data
npx_df <- fread(here("data/olink_10%missing_MinProb_imputed.csv")) # 10% missing samples and proteins were removed before minProb imputation
gcg <- npx_df[, c("eid", "gcg")]
df <- merge(df, gcg, by = "eid", all.x = TRUE)
```

```{r}
# rename T1D and T2D
setnames(df, "adj_probposs_t1dm", "Type1Diabetes")
setnames(df, "adj_probposs_t2dm", "Type2Diabetes")

# Make sure T1D and gestational are removed from controls in T2D
df$adj_dmstatus_bl <- as.numeric(df$adj_dmstatus_bl)
df$adj_dmstatus_bl <-
    factor(df$adj_dmstatus_bl,
        levels = c(0,1,2,3,4,5),
        labels = c(
            "unlikely_diabetes",
            "poss_gestational",
            "prob_t1d",
            "poss_t1d",
            "prob_t2d",
            "poss_t2d"))
df[adj_dmstatus_bl == "poss_gestational" |
       adj_dmstatus_bl == "prob_t1d" |
       adj_dmstatus_bl == "poss_t1d",
   Type2Diabetes := NA] 

df[adj_dmstatus_bl == "poss_gestational" |
       adj_dmstatus_bl == "prob_t2d" |
       adj_dmstatus_bl == "poss_t2d",
   Type1Diabetes := NA]

count(df, adj_dmstatus_bl, Type2Diabetes, Type1Diabetes)

```

Exclude "alcoholics" from PDFF
```{r}
# Import alcoholic/masld df and merge
bs_masld <- fread(here("data/baseline_MASLD.csv"))
df <- merge(df, bs_masld, by = "eid", all.x = TRUE)

# Exclude "alcoholics" from PDFF
count(df, alcoholic, !is.na(PDFF))
df[alcoholic == 1, PDFF := NA]
```


Exclude double mutations (n=3) and rows with NA in all the groups (n=6207)
```{r}
df <- df[!(is.na(trunc) & is.na(cAMP_LoF) & is.na(G40S)),]

double_variant <- which(df$cAMP_LoF == 1 & df$G40S == 1) #gives index's of rows with 1 in cAMP and G40S
df <- df[-double_variant, ]
```


# 2. Linear model with genotype as predictor/independent and Hba1c/BMI/PDFF as outcome

First normalize to SD of each variable
```{r}
# string of variables to loop through
variables <- c("BMI", "Hba1c", "PDFF")

# Calculate SD's
sd_values <- df[, lapply(.SD, sd, na.rm=TRUE), .SDcols = variables]
sd_table <- as.data.frame(t(sd_values))
sd_table$variable <- rownames(sd_table)
#write.xlsx(sd_table, file=here("data/BMI_hba1c_sd_table.xlsx"))

# Loop through the variables and create new columns
for(variable in variables) {
  new_column_name <- paste0(variable, "_SD")
  df[, (new_column_name) := .SD[[variable]] / sd_values[[variable]]]
}
```

Create function to loop over continous variables in linear model, GT as predictor
```{r function_lm}
calculate_coefficients <- function(variables, predictor, data) {
  coef_table <- data.frame()  # Create an empty data frame to store coefficients
  
  for (i in variables) {
    formula_string <- paste(i, "~", predictor, "+age+sex")
    model <- lm(formula_string, data = data)
    
    # Extract coefficients and their standard errors
    coef_summary <- summary(model)$coefficients
    
    # Calculate sample sizes
    sample_sizes <- data %>%
      filter(!is.na(.data[[i]] & !is.na(.data[[predictor]]))) %>%
      group_by(.data[[predictor]]) %>%
      summarise(sample_size = n()) %>%
      pull(sample_size)
    
    # Create a data frame with the p.values for each of the predictors
    coef_data <- data.frame(
      variable = i,
      sample_size = sample_sizes[2],
      GCGR = predictor,
      Estimate = coef_summary[predictor, "Estimate"],
      SE = coef_summary[predictor, "Std. Error"],
      P.value_variable = coef_summary[predictor, "Pr(>|t|)"])
    
    # Append the data frame to the coefficients table
    coef_table <- rbind(coef_table, coef_data)
  }
  
  return(coef_table)
}
```


```{r lm_models}
# string of variables to loop through (continous)
variables <- c("BMI_SD", "Hba1c_SD", "PDFF_SD")

coef_table_G40S <- calculate_coefficients(variables, "G40S", df)
coef_table_cAMP_LoF <- calculate_coefficients(variables, "cAMP_LoF", df)
coef_table_trunc <- calculate_coefficients(variables, "trunc", df)

combined_coef_table <- rbind(coef_table_G40S, coef_table_cAMP_LoF, coef_table_trunc)

combined_coef_table$GCGR[combined_coef_table$GCGR == "trunc"] <- "frameshift"
```

Plot with GT split in color
```{r}
# Remove the "_SD" suffix on the variables
  combined_coef_table$variable <- sub("_SD$", "", combined_coef_table$variable)

p <- ggplot(combined_coef_table, aes(x = Estimate, xmin = Estimate - 1.96 * SE, xmax = Estimate + 1.96 * SE, y = variable, color = GCGR)) +
  geom_point(position = position_dodge(width = 0.8), size=3) +
  geom_errorbarh(position = position_dodge(width = 0.8), height = 0) +
  scale_x_continuous() +
  labs(title = "", x = expression(paste(beta, " in phenotype SD (95% CI)")), y = "") +
    geom_vline(xintercept = 0, linetype = "dashed", color = "gray") +  # Add a dashed line at x = 0
  theme_classic() +
    theme(legend.position = "left",
          axis.line.y = element_blank(),
          axis.ticks.y = element_blank(),
          axis.text.y = element_text(size = 12),
          axis.text.x = element_text(size = 12)) +
    guides(color = guide_legend(reverse = TRUE)) # make legend same order as the data
p
```

```{r tables_for_figure}
# Create the first table
  t1 <- ggplot(data = combined_coef_table) +
    geom_text(aes(y = variable, x = 1, group = GCGR, label = paste0(round(Estimate, digits = 3))), vjust = 0,
              position = position_dodge(width = 0.8)) +
    ggtitle(expression(beta)) +
    xlab(" ") +
    theme_classic(base_size = 13) +
    theme(
      axis.line = element_blank(),
      axis.text.y = element_blank(),
      axis.ticks = element_blank(),
      axis.title.y = element_blank(),
      axis.text.x = element_text(color = "white"),
      plot.title = element_text(size = 12, hjust = 0.5)
    )

 # Create the second table
  t2 <- ggplot(data = combined_coef_table) +
    geom_text(aes(y = variable, x = 1, group = GCGR, 
                  label = paste0("(", round(Estimate - 1.96 * SE, digits = 2), "; ", round(Estimate + 1.96 * SE, digits = 2), ")")), 
              vjust = 0,
              position = position_dodge(width = 0.8)) +
    ggtitle("95% CI") +
    xlab("  ") +
    theme_classic(base_size = 13) +
    theme(
      axis.line = element_blank(),
      axis.text.y = element_blank(),
      axis.ticks = element_blank(),
      axis.title.y = element_blank(),
      axis.text.x = element_text(color = "white"),
      plot.title = element_text(size = 12, hjust = 0.5)
    )

# Create the third table
  t3 <- ggplot(data = combined_coef_table) +
    geom_text(aes(y = variable, x = 1, group = GCGR, 
                  label = paste0(round(P.value_variable, digits = 2))), 
              vjust = 0,
              position = position_dodge(width = 0.8)) +
    ggtitle("p-value") +
    xlab("  ") +
    theme_classic(base_size = 13) +
    theme(
      axis.line = element_blank(),
      axis.text.y = element_blank(),
      axis.ticks = element_blank(),
      axis.title.y = element_blank(),
      axis.text.x = element_text(color = "white"),
      plot.title = element_text(size = 12, hjust = 0.5)
    )
  
  # Create the fourth table
  t4 <- ggplot(data = combined_coef_table) +
    geom_text(aes(y = variable, x = 1, group = GCGR, 
                  label = paste0(sample_size)), 
              vjust = 0,
              position = position_dodge(width = 0.8)) +
    ggtitle("N") +
    xlab("  ") +
    theme_classic(base_size = 13) +
    theme(
      axis.line = element_blank(),
      axis.text.y = element_blank(),
      axis.ticks = element_blank(),
      axis.title.y = element_blank(),
      axis.text.x = element_text(color = "white"),
      plot.title = element_text(size = 12, hjust = 0.5)
    )
```

```{r combine_plot}
# Arrange the plots and tables using grid.arrange
 combined_plot <- grid.arrange(p, t1, t2, t3, t4, widths = c(6, 1, 2, 1, 1))

```

# 4. Logistic model (Firth correction) with T2D/obesity as outcome, GT as predictor
Create function to loop over 0/1 variables in logistic model, GT as predictor. Results is OR
```{r function_glm}
calculate_coefficients_glm <- function(variables, predictor, data) {
  coef_table <- data.frame()  # Create an empty data frame to store coefficients
  
  for (i in variables) {
    formula_string <- paste(i, "~", predictor, "+age+sex")
     model <- logistf(formula_string, data = df)
  
  # Extract coefficients and their standard errors
  coef_summary <- exp(cbind(Odds_Ratio = coef(model), confint(model))) # The coefficients in summary(model) indicate the average change in log odds of disease
  
  # Calculate sample sizes
  sample_sizes <- data %>%
  filter(!is.na(.data[[i]]) & !is.na(.data[[predictor]])) %>%
  group_by(.data[[predictor]], .data[[i]]) %>%
  summarise(sample_size = n()) %>%
  pull(sample_size)

# Check if the length of sample_sizes is 6
if (length(sample_sizes) == 6) {
  # Sum the 4th and 6th values in sample_sizes
  sample_size_genotype_disease <- sample_sizes[4] + sample_sizes[6]
} else {
  # Use the sample size for the genotype with disease as is
  sample_size_genotype_disease <- sample_sizes[4]
}

    # Create a data frame with the p.values for each of the predictors
  coef_data <- data.frame(
    variable = i,
    sample_size_cases = sample_size_genotype_disease,
    GCGR = predictor,
    Odds_ratio = coef_summary[predictor, "Odds_Ratio"],
    CI_low = coef_summary[predictor, "Lower 95%"],
    CI_high = coef_summary[predictor, "Upper 95%"],
    p = model$prob[predictor])
  
  
    # Append the data frame to the coefficients table
  coef_table <- rbind(coef_table, coef_data)
  }
  
  return(coef_table)
}



```

```{r apply_function}
variables <- c("Type2Diabetes", "obesity")
glm_G40S <- calculate_coefficients_glm(variables, "G40S", df)

variables <- c("obesity") # only obesity as too few cases w T2D
glm_cAMP <- calculate_coefficients_glm(variables, "cAMP_LoF", df)
glm_trunc <- calculate_coefficients_glm(variables, "trunc", df)

combined_glm_table <- rbind(glm_G40S, glm_cAMP, glm_trunc)

combined_glm_table$GCGR[combined_glm_table$GCGR == "trunc"] <- "frameshift"
```

Plot with GT split in color
```{r make_plot}
p.glm <- ggplot(combined_glm_table, aes(x = Odds_ratio, xmin = CI_low, xmax = CI_high, y = variable, color = GCGR)) +
  geom_point(position = position_dodge(width = 0.8), size=3) +
  geom_errorbarh(position = position_dodge(width = 0.8), height = 0) +
  scale_x_continuous() +
    #scale_y_discrete(limits = rev(levels(combined_coef_table$variable))) + 
  labs(title = "", x = "OR (95% CI)", y = "") +
    geom_vline(xintercept = 1, linetype = "dashed", color = "gray") +  # Add a dashed line at x = 0
  theme_classic() +
    theme(legend.position = "left",
          axis.line.y = element_blank(),
          axis.ticks.y = element_blank(),
          axis.text.y = element_text(size = 12),
          axis.text.x = element_text(size = 12)) +
    guides(color = guide_legend(reverse = TRUE))
```

```{r tables_for_figure}
# Create the first table
  t1.glm <- ggplot(data = combined_glm_table) +
    geom_text(aes(y = variable, x = 1, group = GCGR, label = paste0(round(Odds_ratio, digits = 2))), vjust = 0,
              position = position_dodge(width = 0.8)) +
    ggtitle("OR") +
    xlab(" ") +
    theme_classic(base_size = 13) +
    theme(
      axis.line = element_blank(),
      axis.text.y = element_blank(),
      axis.ticks = element_blank(),
      axis.title.y = element_blank(),
      axis.text.x = element_text(color = "white"),
      plot.title = element_text(size = 12, hjust = 0.5)
    )

 # Create the second table
  t2.glm <- ggplot(data = combined_glm_table) +
    geom_text(aes(y = variable, x = 1, group = GCGR, 
                  label = paste0("(", round(CI_low, digits = 2), "; ", round(CI_high, digits = 2), ")")), 
              vjust = 0,
              position = position_dodge(width = 0.8)) +
    ggtitle("95% CI") +
    xlab("  ") +
    theme_classic(base_size = 13) +
    theme(
      axis.line = element_blank(),
      axis.text.y = element_blank(),
      axis.ticks = element_blank(),
      axis.title.y = element_blank(),
      axis.text.x = element_text(color = "white"),
      plot.title = element_text(size = 12, hjust = 0.5)
    )

  # Create the third table
  t3.glm <- ggplot(data = combined_glm_table) +
    geom_text(aes(y = variable, x = 1, group = GCGR, 
                  label = paste0(round(p, digits = 2))), 
              vjust = 0,
              position = position_dodge(width = 0.8)) +
    ggtitle("p-value") +
    xlab("  ") +
    theme_classic(base_size = 13) +
    theme(
      axis.line = element_blank(),
      axis.text.y = element_blank(),
      axis.ticks = element_blank(),
      axis.title.y = element_blank(),
      axis.text.x = element_text(color = "white"),
      plot.title = element_text(size = 12, hjust = 0.5)
    )
  
# Create the fourth table
  t4.glm <- ggplot(data = combined_glm_table) +
    geom_text(aes(y = variable, x = 1, group = GCGR, 
                  label = paste0(sample_size_cases)), 
              vjust = 0,
              position = position_dodge(width = 0.8)) +
    ggtitle("N") +
    xlab("  ") +
    theme_classic(base_size = 13) +
    theme(
      axis.line = element_blank(),
      axis.text.y = element_blank(),
      axis.ticks = element_blank(),
      axis.title.y = element_blank(),
      axis.text.x = element_text(color = "white"),
      plot.title = element_text(size = 12, hjust = 0.5)
    )
```

```{r combine_plot}
# Arrange the plots and tables using grid.arrange
 combined_plot.glm <- grid.arrange(p.glm, t1.glm, t2.glm, t3.glm, t4.glm, widths = c(8, 1, 2, 1,1))

```



# 5. Linear model with gcg as outcome, GT as predictor
Update function calculate_coefficints to include BMI
```{r function_lm}
calculate_coefficients <- function(variables, predictor, data) {
  coef_table <- data.frame()  # Create an empty data frame to store coefficients
  
  for (i in variables) {
    formula_string <- paste(i, "~", predictor, "+age+sex+BMI+fasting_time+pCreatinine")
    model <- lm(formula_string, data = data)
    
    # Extract coefficients and their standard errors
    coef_summary <- summary(model)$coefficients
    
    # Calculate sample sizes
    sample_sizes <- data %>%
      filter(!is.na(.data[[i]] & !is.na(.data[[predictor]]))) %>%
      group_by(.data[[predictor]]) %>%
      summarise(sample_size = n()) %>%
      pull(sample_size)
    
    # Create a data frame with the p.values for each of the predictors
    coef_data <- data.frame(
      variable = i,
      sample_size = sample_sizes[2],
      GCGR = predictor,
      Estimate = coef_summary[predictor, "Estimate"],
      SE = coef_summary[predictor, "Std. Error"],
      P.value_variable = coef_summary[predictor, "Pr(>|t|)"])
    
    # Append the data frame to the coefficients table
    coef_table <- rbind(coef_table, coef_data)
  }
  
  return(coef_table)
}
```


```{r}
coef_table_gcg_G40S <- calculate_coefficients("gcg", "G40S", df)
coef_table_gcg_cAMP <- calculate_coefficients("gcg", "cAMP_LoF", df)
coef_table_gcg_trunc <- calculate_coefficients("gcg", "trunc", df)

combined_coef_table_gcg <- rbind(coef_table_gcg_G40S, coef_table_gcg_cAMP, coef_table_gcg_trunc)

combined_coef_table_gcg$GCGR[combined_coef_table_gcg$GCGR == "trunc"] <- "frameshift"
```

Plot with GT split in color
```{r}
p_gcg <- ggplot(combined_coef_table_gcg, aes(x = Estimate, xmin = Estimate - 1.96 * SE, xmax = Estimate + 1.96 * SE, y = variable, color = GCGR)) +
  geom_point(position = position_dodge(width = 0.8), size=3) +
  geom_errorbarh(position = position_dodge(width = 0.8), height = 0) +
      scale_x_continuous() +
    xlim(-1.5, 2) +
     labs(title = "", x = expression(paste(beta, " (95% CI)")), y = "") +
    geom_vline(xintercept = 0, linetype = "dashed", color = "gray") +  # Add a dashed line at x = 0
  theme_classic() +
    theme(legend.position = "left",
          axis.line.y = element_blank(),
          axis.ticks.y = element_blank(),
          axis.text.y = element_text(size = 12),
          axis.text.x = element_text(size = 12)) +
    guides(color = guide_legend(reverse = TRUE))

```


```{r tables_for_figure}
# Create the first table
  t1_gcg <- ggplot(data = combined_coef_table_gcg) +
    geom_text(aes(y = variable, x = 1, group = GCGR, label = paste0(round(Estimate, digits = 3))), vjust = 0,
              position = position_dodge(width = 0.8)) +
    ggtitle(expression(beta)) +
    xlab(" ") +
    theme_classic(base_size = 13) +
    theme(
      axis.line = element_blank(),
      axis.text.y = element_blank(),
      axis.ticks = element_blank(),
      axis.title.y = element_blank(),
      axis.text.x = element_text(color = "white"),
      plot.title = element_text(size = 12, hjust = 0.5)
    )

 # Create the second table
  t2_gcg <- ggplot(data = combined_coef_table_gcg) +
    geom_text(aes(y = variable, x = 1, group = GCGR, 
                  label = paste0("(", round(Estimate - 1.96 * SE, digits = 2), "; ", round(Estimate + 1.96 * SE, digits = 2), ")")), 
              vjust = 0,
              position = position_dodge(width = 0.8)) +
    ggtitle("95% CI") +
    xlab("  ") +
    theme_classic(base_size = 13) +
    theme(
      axis.line = element_blank(),
      axis.text.y = element_blank(),
      axis.ticks = element_blank(),
      axis.title.y = element_blank(),
      axis.text.x = element_text(color = "white"),
      plot.title = element_text(size = 12, hjust = 0.5)
    )

# Create the third table
  t3_gcg <- ggplot(data = combined_coef_table_gcg) +
    geom_text(aes(y = variable, x = 1, group = GCGR, 
                  label = paste0(round(P.value_variable, digits = 2))), 
              vjust = 0,
              position = position_dodge(width = 0.8)) +
    ggtitle("p-value") +
    xlab("  ") +
    theme_classic(base_size = 13) +
    theme(
      axis.line = element_blank(),
      axis.text.y = element_blank(),
      axis.ticks = element_blank(),
      axis.title.y = element_blank(),
      axis.text.x = element_text(color = "white"),
      plot.title = element_text(size = 12, hjust = 0.5)
    )
  
  # Create the fourth table
  t4_gcg <- ggplot(data = combined_coef_table_gcg) +
    geom_text(aes(y = variable, x = 1, group = GCGR, 
                  label = paste0(sample_size)), 
              vjust = 0,
              position = position_dodge(width = 0.8)) +
    ggtitle("N") +
    xlab("  ") +
    theme_classic(base_size = 13) +
    theme(
      axis.line = element_blank(),
      axis.text.y = element_blank(),
      axis.ticks = element_blank(),
      axis.title.y = element_blank(),
      axis.text.x = element_text(color = "white"),
      plot.title = element_text(size = 12, hjust = 0.5)
    )
```


```{r combine_plot}
# Arrange the plots and tables using grid.arrange
 combined_plot <- grid.arrange(p_gcg, t1_gcg, t2_gcg, t3_gcg, t4_gcg, widths = c(6, 1, 2, 1, 1))

```






